---
title: "Covid Data Analysis"
output:
  pdf_document: default
  word_document: default
---

R Libraries
```{r}
library(tidyverse)
library(reticulate)
library(readxl)
py_install("xlrd", pip =TRUE)
library(lubridate)
library(modelr)
library(ggmap)
library(cluster) 
library(factoextra)
library(ggpubr)

```
Importing Data Sets From GitHub
```{r}
USA_Deaths <- read_csv("https://raw.githubusercontent.com/nytimes/covid-19-data/master/us.csv")
USA_States <- read_csv("https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv")
USA_Counties <- read_csv("https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv")
County_Mask <- read_csv("https://raw.githubusercontent.com/nytimes/covid-19-data/master/mask-use/mask-use-by-county.csv")
USA_Counties_Live <- read_csv("https://raw.githubusercontent.com/nytimes/covid-19-data/master/live/us-counties.csv")
Excess_Deaths <- read_csv("https://raw.githubusercontent.com/nytimes/covid-19-data/master/excess-deaths/deaths.csv")

College_Covid <- read_csv("https://raw.githubusercontent.com/nytimes/covid-19-data/master/colleges/colleges.csv")

Geo_Codes <- read_csv("https://raw.githubusercontent.com/Data4Democracy/zip-code-to-county/master/county-fips.csv")

USA_State_Population <- read_csv("https://raw.githubusercontent.com/CivilServiceUSA/us-states/master/data/states.csv")
```


Deaths and Cases Over Time
```{r}

ggplot(data = USA_Deaths, aes(x= date)) +
  geom_line(aes(y=deaths, color = "deaths"))+
  ylab("Deaths")+
  xlab("Month")+
  ggtitle("Deaths Over Time")

ggplot(data = USA_Deaths, aes(x= date)) +
  geom_line(aes(y=cases, color = "cases"))+
  ylab("Cases")+
  xlab("Month")+
  ggtitle("Cases Over Time")
```


Based off of this graph, it appears that cases are growing at a much faster rate then deaths.

```{r}
case_death_ratio <- function(cases,deaths) {
  deaths/cases
}

USA_Deaths %>% mutate( death_ratio = case_death_ratio(cases, deaths)) -> USA_Deaths

ggplot(data = USA_Deaths, aes(x= date)) +
  geom_line(aes(y=death_ratio, color = "Death to Cases Ratio"))+
  ylab("Death to Case Ratio Over Time ")+
  xlab("Month")+
  ggtitle("Death to Case Ratio")

```

Based off of this graph, it appears that deaths per case ratio has decreased over time. 

Masks Usage Analysis Versus Income. Please note the income data is from https://www.indexmundi.com/facts/united-states/quick-facts/illinois/median-household-income#tableTab

```{r}
County_Mask %>% 
  separate(COUNTYFP, into = c('State', 'County'), sep = 2)->County_Mask_State

County_Mask_State %>%
  filter(State == '17') ->Illinois_Data


County_Name <- vector(mode = "character", length = nrow(Illinois_Data))

for (i in 1: nrow(Illinois_Data))
{
  County_Name[i] = ifelse(Illinois_Data$County[i] == "031", "Cook",
                         ifelse(Illinois_Data$County[i] == "097", "Lake", 
                         ifelse(Illinois_Data$County[i] == "043", "DuPage", 
                         ifelse(Illinois_Data$County[i] == "037", "DeKalb", 
                         ifelse(Illinois_Data$County[i] == "089", "Kane",
                         ifelse(Illinois_Data$County[i] == "197", "Will",
                         ifelse(Illinois_Data$County[i] == "111", "McHenry",
                         ifelse(Illinois_Data$County[i] == "003", "Alexander",
                         ifelse(Illinois_Data$County[i] == "121", "Marion",
                         ifelse(Illinois_Data$County[i] == "095", "Knox",
                         ifelse(Illinois_Data$County[i] == "013", "Calhoun",
                         ifelse(Illinois_Data$County[i] == "023", "Clark",
                         ifelse(Illinois_Data$County[i] == "011", "Bureau",
                         ifelse(Illinois_Data$County[i] == "167", "Sangamon",
                         ifelse(Illinois_Data$County[i] == "203", "Woodford",
                         ifelse(Illinois_Data$County[i] == "017", "Cass",
                         ifelse(Illinois_Data$County[i] == "025", "Clay",
                         ifelse(Illinois_Data$County[i] == "051", "Fayette",
                         ifelse(Illinois_Data$County[i] == "061", "Greene",
                         ifelse(Illinois_Data$County[i] == "147", "Piatt",
                         ifelse(Illinois_Data$County[i] == "077", "Jackson",
                         ifelse(Illinois_Data$County[i] == "153", "Pulaski",
                         "Not_Analyzing"))))))))))))))))))))))
}


data.frame(Illinois_Data, County_Name) -> Illinois_Data

Median_Income <- vector(mode = "double", length = nrow(Illinois_Data))

for (i in 1: nrow(Illinois_Data))
{
  Median_Income[i] = ifelse(Illinois_Data$County[i] == "031", 62088,
                         ifelse(Illinois_Data$County[i] == "097", 86244, 
                         ifelse(Illinois_Data$County[i] == "043", 88711, 
                         ifelse(Illinois_Data$County[i] == "037", 61086, 
                         ifelse(Illinois_Data$County[i] == "089", 76912,
                         ifelse(Illinois_Data$County[i] == "197", 83504,
                         ifelse(Illinois_Data$County[i] == "111", 84503,
                         ifelse(Illinois_Data$County[i] == "003", 34709,
                         ifelse(Illinois_Data$County[i] == "121", 46102,
                         ifelse(Illinois_Data$County[i] == "095", 43709,
                         ifelse(Illinois_Data$County[i] == "013", 54392,
                         ifelse(Illinois_Data$County[i] == "023", 54158,
                         ifelse(Illinois_Data$County[i] == "011", 55940,
                         ifelse(Illinois_Data$County[i] == "167", 60466,
                         ifelse(Illinois_Data$County[i] == "203", 71867,
                         ifelse(Illinois_Data$County[i] == "017", 51977,
                         ifelse(Illinois_Data$County[i] == "025", 48691,
                         ifelse(Illinois_Data$County[i] == "051", 44432,
                         ifelse(Illinois_Data$County[i] == "061", 46052,
                         ifelse(Illinois_Data$County[i] == "147", 69271,
                         ifelse(Illinois_Data$County[i] == "077", 37802,
                         ifelse(Illinois_Data$County[i] == "153", 33799,
                         0))))))))))))))))))))))      
                       
}

data.frame(Illinois_Data, Median_Income) -> Illinois_Data

Illinois_Data %>%
  filter(County_Name != "Not_Analyzing")-> Illinois_Data_Summarized
  
ggplot(data = Illinois_Data_Summarized, aes(x = Median_Income, y = NEVER, color = County_Name))+ 
  geom_point()+
  ylab("Percentage Who Answered Never Wears a Mask")+
  xlab("Median Income")+
  ggtitle("Never Wearing a Mask Versus Income Broken Down By County")

ggplot(data = Illinois_Data_Summarized, aes(x = Median_Income, y = ALWAYS, color = County_Name))+ 
  geom_point()+
  ylab("Percentage Who Answered Always Wears a Mask")+
  xlab("Median Income")+
  ggtitle("Always Wearing a Mask Versus Income Broken Down By County")

ggplot(data = Illinois_Data_Summarized, aes(x = Median_Income, y = SOMETIMES, color = County_Name))+ 
  geom_point()+
  ylab("Percentage Who Answered Sometimes Wears a Mask")+
  xlab("Median Income")+
  ggtitle("Sometimes Wearing a Mask Versus Income Broken Down By County")



```
Based off of these graph, it appears that higher income counties are not only more likely to wear a mask, but they have a more adamant about always wearing a mask. Lower income counties are less likely to wear a mask, and have less conviction with their mask usage.  

Linear Regression Used to Analyze How States Performed 

```{r}

USA_States %>% mutate( death_ratio = case_death_ratio(cases, deaths)) -> USA_States

USA_States %>%
  group_by(state)%>%
  summarise(total_cases = sum(cases), total_deaths = sum(deaths), overall_death_ratio = sum(deaths)/sum(cases))-> Summarized_USA_States

Summarized_USA_States %>%
  inner_join(USA_State_Population, by = c("state" = "state"))->Summarized_USA_States

Summarized_USA_States%>%
  arrange(overall_death_ratio)
ggplot(data = Summarized_USA_States, aes(x = total_cases, y = total_deaths, color = state))+ 
  geom_point()+
  ylab("Cases")+
  xlab("Deaths")+
  ggtitle("Death's as a Function of Cases by State")

ggplot(data = Summarized_USA_States, aes(x = population, y = overall_death_ratio, color = state))+ 
  geom_point()+
  ylab("Death Ratio")+
  xlab("Population")+
  ggtitle("Death Ratio as a Function of Population by State")

ggplot(data = Summarized_USA_States, aes(x = population, y = total_deaths, color = state))+ 
  geom_point()+
  ylab("Deaths")+
  xlab("Population")+
  ggtitle("Death's as a Function of Population by State")

ggplot(data = Summarized_USA_States, aes(x = population, y = total_cases, color = state))+ 
  geom_point()+
  ylab("Cases")+
  xlab("Population")+
  ggtitle("Case's as a Function of Population by State")

```




```{r}
lm_model_deaths_vs_cases <- lm(data = Summarized_USA_States, total_deaths ~ total_cases)

summary(lm_model_deaths_vs_cases)

Summarized_USA_States %>%
  add_residuals(lm_model_deaths_vs_cases) %>%
  add_predictions(lm_model_deaths_vs_cases) -> Summarized_USA_States_LM_deaths_vs_cases

Summarized_USA_States_LM_deaths_vs_cases %>%
  select(state, total_deaths, total_cases, resid, pred)%>%
  arrange(resid)%>%
  head(8)->Summarized_USA_States_LM_deaths_vs_cases_top8


ggplot(data=Summarized_USA_States_LM_deaths_vs_cases_top8, aes(x=state, y=resid, fill = state)) +
  geom_bar(stat="identity")+
  ylab("Residual Deaths")+
  xlab("")+
  ggtitle("Top 8 States for Death Residuals (Cases)")

Summarized_USA_States_LM_deaths_vs_cases %>%
  select(state, total_deaths, total_cases, resid, pred)%>%
  arrange(-resid)%>%
  head(8)->Summarized_USA_States_LM_deaths_vs_cases_bottom8

ggplot(data=Summarized_USA_States_LM_deaths_vs_cases_bottom8, aes(x=state, y=resid, fill = state)) +
  geom_bar(stat="identity")+
  ylab("Residual Deaths")+
  xlab("")+
  ggtitle("Bottom 8 States for Death Residuals (Cases)")

Summarized_USA_States_LM_deaths_vs_cases

```
```{r}
lm_model_deathratio_vs_population <- lm(data = Summarized_USA_States, overall_death_ratio ~ population)

summary(lm_model_deathratio_vs_population) 

Summarized_USA_States %>%
  add_residuals(lm_model_deathratio_vs_population) %>%
  add_predictions(lm_model_deathratio_vs_population) -> Summarized_USA_States_LM_deathratio_vs_population

Summarized_USA_States_LM_deathratio_vs_population %>%
  select(state, total_deaths, overall_death_ratio, resid, pred)%>%
  arrange(resid)%>%
  head(8)->Summarized_USA_States_LM_deaths_vs_cases_top8


Summarized_USA_States_LM_deaths_vs_cases_top8

Summarized_USA_States_LM_deathratio_vs_population %>%
  select(state, total_deaths, overall_death_ratio, resid, pred)%>%
  arrange(-resid)%>%
  head(8)->Summarized_USA_States_LM_deaths_vs_cases_bottom8

Summarized_USA_States_LM_deaths_vs_cases_bottom8

ggplot(data=Summarized_USA_States_LM_deaths_vs_cases_top8, aes(x=state, y=resid, fill = state)) +
  geom_bar(stat="identity")+
  ylab("Residual Death Ratio")+
  xlab("")+
  ggtitle("Top 8 Performing States for Death Ratio Residuals")

ggplot(data=Summarized_USA_States_LM_deaths_vs_cases_bottom8, aes(x=state, y=resid, fill = state)) +
  geom_bar(stat="identity")+
  ylab("Residual Death Ratio")+
  xlab("")+
  ggtitle("Bottom 8 Performing States for Death Ratio Residuals")

```

```{r}
lm_model_deathS_vs_population <- lm(data = Summarized_USA_States, total_deaths ~ population)

lm(data = Summarized_USA_States, total_deaths ~ population)

summary(lm_model_deathS_vs_population) 

Summarized_USA_States %>%
  add_residuals(lm_model_deathS_vs_population) %>%
  add_predictions(lm_model_deathS_vs_population) -> Summarized_USA_States_LM_deaths_vs_population

Summarized_USA_States_LM_deaths_vs_population %>%
  select(state, total_deaths, overall_death_ratio, resid, pred)%>%
  arrange(resid)%>%
  head(8)->Summarized_USA_States_LM_deaths_vs_population_top8

Summarized_USA_States_LM_deaths_vs_population_top8

ggplot(data=Summarized_USA_States_LM_deaths_vs_population_top8, aes(x=state, y=resid, fill = state)) +
  geom_bar(stat="identity")+
  ylab("Residual Deaths")+
  xlab("")+
  ggtitle("Top 8 Performing States for Death Total Residuals (Population)")

Summarized_USA_States_LM_deaths_vs_population %>%
  select(state, total_deaths, overall_death_ratio, resid, pred)%>%
  arrange(-resid)%>%
  head(8)->Summarized_USA_States_LM_deaths_vs_population_bottom8

Summarized_USA_States_LM_deaths_vs_population_bottom8

ggplot(data=Summarized_USA_States_LM_deaths_vs_population_bottom8, aes(x=state, y=resid, fill = state)) +
  geom_bar(stat="identity")+
  ylab("Residual Deaths")+
  xlab("")+
  ggtitle("Bottom 8 Performing States for Death Total Residuals (Population)")


```

```{r}
lm_model_cases_vs_population <- lm(data = Summarized_USA_States, total_cases ~ population)

summary(lm_model_cases_vs_population)

Summarized_USA_States %>%
  add_residuals(lm_model_cases_vs_population) %>%
  add_predictions(lm_model_cases_vs_population) -> Summarized_USA_States_LM_cases_vs_population

Summarized_USA_States_LM_cases_vs_population %>%
  select(state, total_deaths, overall_death_ratio, resid, pred)%>%
  arrange(resid)%>%
  head(8)->Summarized_USA_States_LM_cases_vs_population_top8

Summarized_USA_States_LM_cases_vs_population_top8

ggplot(data=Summarized_USA_States_LM_cases_vs_population_top8, aes(x=state, y=resid, fill = state)) +
  geom_bar(stat="identity")+
  ylab("Residual Cases")+
  xlab("")+
  ggtitle("Top 8 Performing States for Case Total Residuals")

Summarized_USA_States_LM_cases_vs_population %>%
  select(state, total_deaths, overall_death_ratio, resid, pred)%>%
  arrange(-resid)%>%
  head(8)->Summarized_USA_States_LM_cases_vs_population_bottom8

Summarized_USA_States_LM_cases_vs_population_bottom8

ggplot(data=Summarized_USA_States_LM_cases_vs_population_bottom8, aes(x=state, y=resid, fill = state)) +
  geom_bar(stat="identity")+
  ylab("Residual Cases")+
  xlab("")+
  ggtitle("Bottom 8 Performing States for Case Total Residuals")
```

